Response to reviewers and editorial comments for "Spatio-temporal estimates of HIV risk group proportions for adolescent

### 2023-01-31 src/docs_paper/paper.Rmd
Final updates to review response

The Global AIDS Strategy 2021-2026 [@unaids2021global], adopted by the United Nations (UN) General Assembly in June 2021, proposed stratifying packages of HIV prevention provided to AGYW based on both local population-level HIV incidence and individual-level sexual risk behaviour to promote more efficient priorisation of prevention services [@anderson2014maximising; @mcgillen2016optimum; @kelly2018global; @jia2022risk].
Four prioritisation strata (Supplementary Table B.1) were defined based on: (1) subnational annual incidence (<0.3\%, 0.3-1.0\%, 1.0-3.0\% and >3.0\%), and (2) self-reported high-risk behaviour or recent STI infection.
<!-- NB: Check again about the recent STI infection aspect. -->
The strategy encourages programmes to  define targets for the proportion of AGYW to be reached with a range of interventions (Supplementary Table B.2) based upon this prioritisation strata [@unaids2021global].
All AGYW are recommended to have access to a basic package of HIV prevention, while those with with high risk behaviours in moderate incidence settings and all AGYW in very high incidence settings are recommended to have access to enhanced intervention packages [@joint2016hiv; @birdthistle2018evaluating].
<!-- Could add "basic package of HIV prevention consisting of... -->
These interventions may include STI screening and treatment, access to pre-exposure prophylaxis (PrEP), access to post-exposure prophylaxis (PEP), comprehensive sexuality education, and economic empowerment.

@@

We have the least confidence in our estimates for the FSW risk group.
As well as having the smallest sample sizes, our transactional sex estimates do not overcome the difficulties of sampling hard to reach groups.
We inherent any limitations of the national FSW estimates [@stevens2022estimating] which we adjust our estimates of transactional sex to match.
<!-- (As presented by Oli's paper) -->
Furthermore, we do not consider seasonal migration patterns, which may particularly affect FSW size.
More generally, we did not aim to include predictive covariates.
Many covariates that might be predictive of risk group proportions (for example sociodemographic characteristics, education, local economic activity, cultural and religious norms and attitudes) are themselves difficult to measure spatially, and previous efforts have only led to modest improvements in model performance.
Identifying measurable correlates of risk, or particular settings in which time-concentrated HIV risk occurs, is an important area for further research to improve risk priorisation and precision HIV programme delivery.

@@

## Funding

AH was supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
AH, VKN, KAR, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661, OPP1164897).
TMW was supported by Imperial College London (President's PhD Scholarship).
SRF was supported by the EPSRC (EP/V002910/2).
JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664). 
LZ, IW, MM and CB are employed by UNAIDS.
A first version of this work was funded by UNAIDS.
This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.

For the purpose of Open Access, the authors have applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.

## Disclaimer

The findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies.

## Author contributions

* Conceptualisation: JWE, KAR, CB
* Data curation: AH, KAR, VKN, RE, OS, JWE
* Formal analysis: AH
* Funding acquisition: JWE
* Investigation: AH
* Methodology: AH, KAR, VKN, TMW, SRF, JWE
* Project administration: JWE, KAR, CB
* Software: AH, KAR
* Supervision: JWE, SRF
* Visualisation: AH
* Writing - Original draft: AH
* Writing - Review and Editing: AH, KAR, VKN, OS, KMJ, RE, LZ, IW, MM, CB, SRF, JWE

>>>>

The Global AIDS Strategy 2021-2026 [@unaids2021global], adopted by the United Nations (UN) General Assembly in June 2021, proposed stratifying packages of HIV prevention provided to AGYW based on both local population-level HIV incidence and individual-level sexual risk behaviour to promote more efficient prioritisation of prevention services [@anderson2014maximising; @mcgillen2016optimum; @kelly2018global; @jia2022risk].
Four prioritisation strata (Supplementary Table B.1) were defined based on: (1) subnational annual incidence (<0.3\%, 0.3-1.0\%, 1.0-3.0\% and >3.0\%), and (2) self-reported high-risk behaviour or recent STI infection.
<!-- NB: Check again about the recent STI infection aspect. -->
The strategy encourages programmes to  define targets for the proportion of AGYW to be reached with a range of interventions (Supplementary Table B.2) based upon this prioritisation strata [@unaids2021global].
All AGYW are recommended to have access to a basic package of HIV prevention, while those with with high risk behaviours in moderate incidence settings and all AGYW in very high incidence settings are recommended to have access to enhanced intervention packages [@joint2016hiv; @birdthistle2018evaluating].
<!-- Could add "basic package of HIV prevention consisting of... -->
These interventions may include STI screening and treatment, access to pre-exposure prophylaxis (PrEP), access to post-exposure prophylaxis (PEP), comprehensive sexuality education, and economic empowerment.

@@

We have the least confidence in our estimates for the FSW risk group.
As well as having the smallest sample sizes, our transactional sex estimates do not overcome the difficulties of sampling hard to reach groups.
We inherent any limitations of the national FSW estimates [@stevens2022estimating] which we adjust our estimates of transactional sex to match.
<!-- (As presented by Oli's paper) -->
Furthermore, we do not consider seasonal migration patterns, which may particularly affect FSW size.
More generally, we did not consider covariates potentially predictive of risk group proportions (such as sociodemographic characteristics, education, local economic activity, cultural and religious norms and attitudes), which are typically difficult to measure spatially.
Identifying measurable correlates of risk, or particular settings in which time-concentrated HIV risk occurs, is an important area for further research to improve risk prioritisation and precision HIV programme delivery.

@@

<!-- ## Funding and disclaimer -->

<!-- AH was supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1). -->
<!-- AH, VKN, KAR, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661, OPP1164897). -->
<!-- TMW was supported by Imperial College London (President's PhD Scholarship). -->
<!-- SRF was supported by the EPSRC (EP/V002910/2). -->
<!-- JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).  -->
<!-- This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union. -->

<!-- LZ, IW, MM, and CB are employed by UNAIDS, who also partially funded this study. -->
<!-- LZ, IW, MM, and CB contributed to the design of the study, revision of the manuscript, and approved the final manuscript for submission. -->
<!-- Findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies. -->

<!-- ## Author contributions -->

<!-- * Conceptualisation: JWE, KAR, CB -->
<!-- * Data curation: AH, KAR, VKN, RE, OS, JWE -->
<!-- * Formal analysis: AH -->
<!-- * Funding acquisition: JWE -->
<!-- * Investigation: AH -->
<!-- * Methodology: AH, KAR, VKN, TMW, SRF, JWE -->
<!-- * Project administration: JWE, KAR, CB -->
<!-- * Software: AH, KAR -->
<!-- * Supervision: JWE, SRF -->
<!-- * Visualisation: AH -->
<!-- * Writing - Original draft: AH -->
<!-- * Writing - Review and Editing: AH, KAR, VKN, OS, KMJ, RE, LZ, IW, MM, CB, SRF, JWE -->

### 2023-01-31 src/docs_review/review.Rmd
Final updates to review response

We agree that identifying predictive and measureable predictors of risk group proportions, or high risk locations, is an important area for further research.
We have commented on this in the discussion section:

> More generally, we did not aim to include predictive covariates. Many covariates that might be predictive of risk group proportions (for example sociodemographic characteristics, education, local economic activity, cultural and religious norms and attitudes) are themselves difficult to measure spatially, and previous efforts have only led to modest improvements in model performance. Identifying measurable correlates of risk, or particular settings in which time-concentrated HIV risk occurs, is an important area for further research to improve risk priorisation and precision HIV programme delivery.

@@

Response here.
Requested that coauthors send their COIs.

@@

Response here.

@@

Response here.

@@

Response here.

@@

Response here.

>>>>

We agree that identifying predictive and measurable predictors of risk group proportions, or high risk locations, is an important area for further research.
We have commented on this in the discussion section:

> More generally, we did not consider covariates potentially predictive of risk group proportions (such as sociodemographic characteristics, education, local economic activity, cultural and religious norms and attitudes), which are typically difficult to measure spatially. Identifying measurable correlates of risk, or particular settings in which time-concentrated HIV risk occurs, is an important area for further research to improve risk prioritisation and precision HIV programme delivery.

@@

The competing interests' statement has been updated to the following:

> I have read the journal's policy and the authors of this manuscript have the following competing interests: KAR was supported through a consultancy via UNAIDS to conduct an early version of this work;  LZ, IW, MM, and CB are employees of UNAIDS; JWE is a member of the editorial board for PLOS Global Public Health; other authors have declared that no competing interests exist.

@@

The financial disclosure statement has been amended to clarify the role of the funders as following:

> AH was supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1). AH, VKN, KAR, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661, OPP1164897). TMW was supported by Imperial College London (President's PhD Scholarship). SF was supported by the EPSRC (EP/V002910/2). JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664). LZ, IW, MM and CB are employed by UNAIDS. This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth & Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.

> LZ, IW, MM, and CB are employed by UNAIDS, who also partially funded this study.
LZ, IW, MM, and CB contributed to the design of the study, revision of the manuscript, and approved the final manuscript for submission.
Findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies.

@@

The financial disclosure fields and funding information tab have now been aligned.

@@

Separate figure files have been provided as requested.

@@

We attach an email from UNAIDS confirming approval for use of these subnational boundary shapefiles.
